---
title: "$t$ Distribution"
author: "Dr. Dogucu"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      
---

layout: true
  
<div class="my-header"></div>

<div class="my-footer"> 
 Copyright &copy; <a href="https://mdogucu.ics.uci.edu">Dr. Mine Dogucu</a>. <a href="https://creativecommons.org/licenses/by-nc-sa/4.0/">CC BY-NC-SA 4.0</a></div> 


---
class: middle

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = "center")
library(tidyverse)
options(scipen=0)
theme_set(theme_gray(base_size = 18))

set.seed(84735)

plot_norm_x <- function(value, mean, sd, end = 3){
  data_norm <-  data.frame(x = c(mean-(end*sd), mean+(end*sd)))
  
  data_line <- data.frame(x = value, y = dnorm(value, mean, sd), 
                          label = paste("x = ", value))
  ggplot(data_norm, 
         aes(x)) +
    stat_function(fun = dnorm, n = 1001, args = list(mean = mean, sd = sd)) + 
    ylab("") +
    geom_segment(data = data_line, aes(x = value, y = 0, xend = x, yend = y), color = "#e56646") +
    geom_text(data = data_line, aes(x = x, y = 0, label = label), color = "#e56646") +
    labs(title =  bquote("N("*mu == .(mean) ~ "," ~ sigma == .(sd)*")"))
}



plot_norm_z <- function(value, mean, sd, end = 3){
  
  data_norm <-  data.frame(x = c(mean-(end*sd), mean+(end*sd)))
  data_z <- data.frame(x = c(-end,end))

  
  data_line_z <- data.frame(z = (value-mean)/sd, y = dnorm((value-mean)/sd), 
                            label = paste("z = ", (value-mean)/sd))
  
ggplot(data_z, 
               aes(x)) +
    stat_function(fun = dnorm, n = 1001, args = list(mean = 0, sd = 1)) + 
    xlab("z")+
    ylab("") +
    geom_segment(data = data_line_z, aes(x = z, y = 0, xend = z, yend = y), color = "#e56646") +
    geom_text(data = data_line_z, aes(x = z, y = 0, label = label), color = "#e56646") +
  labs(title =  bquote("N("*mu == 0 ~ "," ~ sigma == 1*")"))

  
}


plot_norm_x_prob <- function(value, mean, sd, end = 3, lower = TRUE){
  
  x <- seq(mean-(end*sd),mean+(end*sd), 0.01)
  
  y <- dnorm(x, mean = mean, sd = sd)
  
  data <- data.frame(x = x, y = y)
  
  data_line <- data.frame(x = value, y = dnorm(value, mean, sd), 
                          label = paste("x = ", value))
  
  if (lower == TRUE){
    p <- qplot(x, y, data = data, 
          geom = "line") +
      geom_ribbon(data = subset(data, x > (mean-(end*sd)) & x < value),
                  aes(ymax = y), 
                  ymin = 0,
                  fill = "#e56646", 
                  colour = NA, 
                  alpha = 0.5) +
      ylab("") 

  }
  
  if (lower == FALSE){
    p <- qplot(x, y, data = data, 
          geom = "line") +
      geom_ribbon(data = subset(data, x < (mean+(end*sd)) & x > value),
                  aes(ymax = y), 
                  ymin = 0,
                  fill = "#e56646", 
                  colour = NA, 
                  alpha = 0.5) +
      ylab("")
  }
  
p
}

```

## Uses of Normal Distribution

- SAT scores are normally distributed we were able to use normal distribution.

--

- Baby weights are normally distributed we were able to use normal distribution.

--

- According to Central Limit Theorem, sample proportions are approximately normally distributed (if certain conditions are met).

--

- According to Central Limit Theorem, difference of two proportions are approximately normally distributed (if certain conditions are met).

---

class: middle

## What about Single Mean?


- According to Central Limit Theorem, sample means are approximately normally distributed (if certain conditions are met).

---

class: middle


## Conditions

We can rely on Central Limit Theorem in order to make inference for a single mean if the following conditions have been met 

__Independence__ The sample data must be independent.

__Normality__ If the sample size is small, then the sample data must come from a population with a normal distribution. 

---

class: middle


## Using CLT for Single Mean

If the conditions are met then

$\bar x \sim \text{approximately }N(mean = \mu, sd = \frac{\sigma}{\sqrt{n}})$

We have a problem! We do not know $\sigma$ so we may not calculate the standard error easily. 

---

class: middle

## How did we handle this problem before?

When making inference for single proportion and difference of two proportions

$p \sim \text{approximately } N(mean = \pi, sd = \sqrt{\frac{\pi(1-\pi)}{n}})$

$(p_1 - \hat p_2) \sim \text{approximately } N((mean = \pi_1 - \pi_2), sd = \sqrt{\frac{\pi_1(1-\pi_1)}{n_1} + \frac{\pi_2(1-\pi_2)}{n_2}})$

we did not know the population proportions. How did we deal with this problem when calculating standard error?

---

class: middle

## How did we handle this problem before?

We used the sample proportions. 

---

class: middle 

## How can we handle our current problem?

Since we do not know $\sigma$, the population standard deviation, we will instead use $s$, the sample standard deviation.

Using $s$ instead of $\sigma$ creates a problem in estimation.

Rather than using Normal distribution, we will now rely on $t$-distribution.

---

## t-distribution

```{r warning = FALSE}
ggplot(data = data.frame(x = c(-3, 3)), aes(x)) +
  stat_function(fun = dt, n = 101, args = list(df = 2)) + 
  scale_y_continuous(breaks = NULL) +
  ylab("f(x)") +
  theme(text = element_text(size = 20))

```

---

## t-distribution vs Normal Distribution

```{r warning = FALSE}

ggplot(data = data.frame(x = c(-3, 3)), aes(x)) +
  stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = 1), aes(color = "z"))  +
  stat_function(fun = dt, n = 101, args = list(df = 1), aes(color = "t")) + 
  ylab("f(x)") +
  scale_y_continuous(breaks = NULL) +
  scale_colour_manual(name="Distribution",
  values=c(z="red", t="blue")) +
  theme(text = element_text(size = 20))


```

---

class: middle


## t-distribution

In $t$-distribution, we will use one parameter only which is named degrees of freedom. 

$df = n - 1$

---

## t-distribution with different df

```{r warning = FALSE}

ggplot(data = data.frame(x = c(-3, 3)), aes(x)) +
  stat_function(fun = dt, n = 101, args = list(df = 1), aes(color = "df1")) + 
  stat_function(fun = dt, n = 101, args = list(df = 2), aes(color = "df2")) + 
  stat_function(fun = dt, n = 101, args = list(df = 4), aes(color = "df4")) + 
  stat_function(fun = dt, n = 101, args = list(df = 7), aes(color = "df7")) + 
  stat_function(fun = dt, n = 101, args = list(df = 10), aes(color = "df9")) + 
  ylab("f(x)") +
  scale_y_continuous(breaks = NULL) +
  scale_colour_manual(name="df",
  values=c(df1 = "red", df2="blue", df4 = "green", df7 = "black", df9 = "orange")) +
  theme(text = element_text(size = 20))


```

## t-distribution vs Normal Distribution

Note that as degrees of freedom increases t-distribution approaches a normal distribution. 
---
## t-distribution using R

```{r fig.height=3}
x <- seq(-4,4, 0.01)
y <- dt(x, df = 1)
data <- data.frame(x = x, y = y)
qplot(x, y, data = data, 
      geom = "line")+
  geom_ribbon(data = subset(data, x >- 4 & x < -2),
              aes(ymax = y), ymin = 0,
              fill = "#e56646", colour = NA, alpha = 0.5)+
  scale_y_continuous(limits = c(0, .4)) +
  ylab("f(x)") 

```


What proportion of values are below -2 in this distribution? 

$X \sim t_{df = 1} \text{ then }P(x \leq -2)$

```{r echo = TRUE}
pt(-2, df = 1) 
```

---

## t-distribution using R

```{r fig.height=3}
x <- seq(-4,4, 0.01)
y <- dt(x, df = 1)
data <- data.frame(x = x, y = y)
qplot(x, y, data = data, 
      geom = "line")+
  geom_ribbon(data = subset(data, x >- 4 & x < -2),
              aes(ymax = y), ymin = 0,
              fill = "#e56646", colour = NA, alpha = 0.5)+
  scale_y_continuous(limits = c(0, .4)) +
  ylab("f(x)") 

```


Below which value does the 14.75836% of the data lie?

```{r echo = TRUE}
qt(0.1475836, df = 1) #
```

---

## Where does the middle 95% of the data lie when $df = 14$? 

```{r fig.height=3}
x <- seq(-4,4, 0.01)
y <- dt(x, df = 14)
data <- data.frame(x = x, y = y)
qplot(x, y, data = data, 
      geom = "line")+
  geom_ribbon(data = subset(data, x >-2.144787 & x < 2.144787),
              aes(ymax = y), ymin = 0,
              fill = "#e56646", colour = NA, alpha = 0.5) +
  scale_y_continuous(limits = c(0, .4)) +
  ylab("f(x)") +
  theme(text = element_text(size = 20))

```

```{r echo = TRUE}
qt(0.025, df = 14)
```

95% of the data lies between -2.144787 and 2.144787

---

## Where does the middle 90% of the data lie when $df = 149$? 

```{r fig.height=3}
x <- seq(-4,4, 0.01)
y <- dt(x, df = 149)
data <- data.frame(x = x, y = y)
qplot(x, y, data = data, 
      geom = "line")+
  geom_ribbon(data = subset(data, x > -1.655145 & x < 1.655145),
              aes(ymax = y), ymin = 0,
              fill = "#e56646", colour = NA, alpha = 0.5) +
  scale_y_continuous(limits = c(0, .4)) +
  ylab("f(x)") +
  theme(text = element_text(size = 20))

```

```{r echo = TRUE}
qt(0.05, df = 149)
```

90% of the data lies between `r qt(0.05, df = 149)` and `r qt(0.05, df = 149)*-1`



