---
title: "Binomial Distribution"
author: "Dr. Dogucu"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      
---

layout: true
  
<div class="my-header"></div>

<div class="my-footer"> 
 Copyright &copy; <a href="https://mdogucu.ics.uci.edu">Dr. Mine Dogucu</a>. <a href="https://creativecommons.org/licenses/by-nc-sa/4.0/">CC BY-NC-SA 4.0</a></div> 

---

## Conditions

- The trials have to be independent from each other.
- The probability of success has to be the same for each trial.
- The number of trials is fixed. 
- Each trial outcome is either a success or a failure.

---

## Binomial Distribution

The random variable X represents the number of successes in $n$ trials where in independent trial the probability of success is $p$.

--

$X\sim \text{Binomial}(n, p)$

--

$P(X = x) = f(x) = {n \choose x}p^{x} (1-p)^{n-x}$

--

$S = \{0,1,2...,n\}$

--

$E(X) = np$

--

$Var(X) = np(1-p)$


---

__Example__

A vet has been assigned to work at a farm where 70% of the animals have been infected by avian influenza. The vet selects 10 random animals to inspect. What is the probability that 3 of the  selected animals are infected? 

--

$n = 10$, $x = 3$, $p = 0.70$

--

$P(X = 3) = f(3) = {10 \choose 3}0.70^{3} (1-0.70)^{10-3}$

--

$P(X = 3)= \frac{10!}{3!7!}0.70^30.30^7=0.009001692$

--
```{r}
dbinom(x = 3, size = 10, prob = 0.70)
```


---

class: center middle

## pmf when $n = 10$ and $p = 0.70$


```{r echo = FALSE, message = FALSE, fig.align="center", fig.height = 5}
library(tidyverse)
theme_set(theme_gray(base_size = 22))

x <- seq(0,10)
y <- dbinom(x, size = 10 , prob = 0.70)

data <- data.frame(x = x,
                   y = y)

data %>% 
  ggplot() +
  aes(x = x, y = y) +
  geom_segment(data = data, 
               aes(x = x, y = 0, 
                   xend = x, yend = y)) +
  labs(x = expression(paste("x")),
       y = expression(paste("f(x)"))) +
      scale_x_continuous(breaks=seq(0,10))


```

---

## Cumulative Probability Function

What is the probability that at most three of the ten selected animals are infected? 

$P(X \leq 3) = P(X = 0) + P(X = 1) + P(X = 2) + P(X = 3)$

--

$\frac{10!}{10!0!}0.70^{0}0.30^10 + \frac{10!}{9!1!}0.70^{1}0.30^9 + \frac{10!}{8!2!}0.70^{2}0.30^8 + \frac{10!}{7!3!}0.70^{3}0.30^7$

--


```{r}
dbinom(x = 0, size = 10, prob = 0.70) +
  dbinom(x = 1, size = 10, prob = 0.70) +
  dbinom(x = 2, size = 10, prob = 0.70) +
  dbinom(x = 3, size = 10, prob = 0.70)

```


---

## Cumulative Probability Function

$P(X \leq 3)$

```{r}
pbinom(q = 3, size = 10, prob = 0.70)
```

---

## Expected Value 

What is the expected value of number of infected animals in 10 selected animals?

--

$E(X) = np$  
$E(X) = 10\times0.70=7$


--

## Variance

What is the variance of number of infected animals in 10 selected animals?

--

$Var(X) = np(1-p)$

$Var(X) = 10\times0.7(1-0.7) = 2.1$
