---
title: "Descriptive Statistics in R"
author: "Dr. Dogucu"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      
---

layout: true
  
<div class="my-header"></div>

<div class="my-footer"> 
 Copyright &copy; <a href="https://mdogucu.ics.uci.edu">Dr. Mine Dogucu</a>. <a href="https://creativecommons.org/licenses/by-nc-sa/4.0/">CC BY-NC-SA 4.0</a></div> 

---
```{r echo = FALSE, message = FALSE}
library(titanic)
library(tidyverse)
titanic <- titanic_train %>% 
  mutate(Survived = as.logical(Survived),
         Sex = as.factor(Sex),
         Embarked = as.factor(Embarked))
  

```

---

## Data

```{r}
glimpse(titanic)
```

---

## Outline

- Summarizing Categorical Data in R


- Summarizing Numerical Data in R

We will only summarize a _single_ variable at a time but later we will talk about relationship between two variables.

---

## Summarizing Categorical Data

__Counts__ or __proportions__ are used to summarize categorical variables. 

They are often summarized with a frequency table.


---

## Summarizing Categorical Data in R

```{r}
count(titanic, Survived)
```

There are 342 people who survived and 549 who did not. 

---

## Summarizing Categorical Data in R


There are total 549 + 342 people. We can use this to calculate the proportions

```{r}
549/(549+342)

342/(549+342)
```


---

## Mean in R

```{r}
summarize(titanic, mean(Fare))
```
---

## Median in R

```{r}
summarize(titanic, mean(Fare))
```

---


## Mode in R 

```{r}
count(titanic, Fare, sort = TRUE)
```

---

## Quartiles

You can specify quartiles (Q3)

```{r}
summarize(titanic, quantile(Fare, 0.75))
```

---
## Standard Deviation and Variance

```{r}
summarize(titanic, sd(Fare))
```

```{r}
summarize(titanic, var(Fare))
```
