---
title: "Discrete Random Variables Review"
author: "Dr. Dogucu"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      
---

layout: true
  
<div class="my-header"></div>

<div class="my-footer"> 
 Copyright &copy; <a href="https://mdogucu.ics.uci.edu">Dr. Mine Dogucu</a>. <a href="https://creativecommons.org/licenses/by-nc-sa/4.0/">CC BY-NC-SA 4.0</a></div> 


---

A **discrete random** variable has a countable number of possible numeric outcomes.

--

Probability mass function (pmf):

$P(X = x) = f(x)$

--

Cumulative distribution function (cdf):

$P(X \leq x) = F(x)$

--

$E(X) = \sum_{S} x f(x)$

--

$Var(X)= E(X^2) - [E(X)]^2$


---


## Bernoulli Distribution

If X is a random variable that takes value 1 with probability of success $p$ and 0 with probability $1-p$, then X follows a Bernoulli distribution.


$X \sim \text{Bernoulli} (p)$

--

$E(X) = \mu = p$

--

$Var(X)=\sigma^2 = p(1-p)$

---


## Geometric Distribution

Let X be the number of failures needed __before__ the first success is observed in independent trials. $X$ follows a geometric distribution

.pull-left[

$S = \{0, 1, 2, 3, 4, ...\}$

$X \sim \text{Geometric} (p)$

$f(x) = (1-p)^x(p)$  

$E(X)=\frac{1-p}{p}$

$Var(X) = \frac{1-p}{p^2}$

]

.pull-right[

```{r eval = FALSE}
dgeom(x, prob) #pmf
```



```{r eval = FALSE}
pgeom(q, prob) #cdf
```

]

---

## Binomial Distribution

The random variable X represents the number of successes in $n$ trials where in independent trial the probability of success is $p$.

.pull-left[

$X\sim \text{Binomial}(n, p)$

$P(X = x) = f(x) = {n \choose x}p^{x} (1-p)^{n-x}$


$E(X) = np$  


$Var(X) = np(1-p)$

]

.pull-right[

```{r eval = FALSE}
dbinom(x, size, prob) #pmf
```


```{r eval = FALSE}
pbinom(q, size, prob) #cdf
```

]

---

## Poisson Distribution

Let $X$ represent the number of occurrences of an event within a  __fixed__ time or space.

.pull-left[

$X \sim Poisson (\lambda)$


$P(X = x) = f(x) =\frac{\lambda^x}{x!} e^{-\lambda}$ 



$E(X) = Var(X) = \lambda$

]

.pull-right[

```{r eval = FALSE}
dpois(x, lambda) #pmf
```


```{r eval = FALSE}
ppois(q, lambda) #cdf
```

]