---
title: "Confidence Interval for Difference of Two Means"
author: "Dr. Dogucu"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      
---

layout: true
  
<div class="my-header"></div>

<div class="my-footer"> 
 Copyright &copy; <a href="https://mdogucu.ics.uci.edu">Dr. Mine Dogucu</a>. <a href="https://creativecommons.org/licenses/by-nc-sa/4.0/">CC BY-NC-SA 4.0</a></div> 


---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = "center")
library(tidyverse)
options(scipen=0)
theme_set(theme_gray(base_size = 18))

```




## Confidence Interval Review


Confidence Interval = point estimate $\pm$ critical value $\times$ standard error of the estimate

|                               | point estimate        | critical value | standard error of the estimate                    |
|-------------------------------|-----------------------|:--------------:|---------------------------------------------------|
| single proportion             | $p$                   | z*             | $\sqrt{\frac{p(1-p)}{n}}$                         |
| difference of two proportions | $p_1-p_2$             | z*             | $\sqrt{\frac{p_1(1-p_1)}{n}+\frac{p_2(1-p_2)}{n}}$ |
| single mean                   | $\bar x$              | $t*_{df}$      | $\sqrt{\frac{s^2}{n}}$                            |
| difference of two means       | $\bar x_1 - \bar x_2$ | $t*_{df}$      | $\sqrt{\frac{s_1^2}{n_1} + \frac{s_2^2}{n_2}}$    |

---
class: middle

## Conditions

1. Independence: Within each group data have to be independent from each other. The two groups have to be independent from one another.

2. Normality: We check for normality for each group.
---


class: center middle


## Confidence Interval for Difference of Two Means


$\bar x_1 - \bar x_2 \pm t^*_{df} \times \sqrt{\frac{s_1^2}{n_1} + \frac{s_2^2}{n_2}}$

---

class: middle

## Example 
Onnasch, L., & Roesler, E. (2019). Anthropomorphizing Robots: **The Effect of Framing in Human-Robot Collaboration.** Proceedings of the Human Factors and Ergonomics Society Annual Meeting, 63(1), 1311–1315. https://doi.org/10.1177/1071181319631209

--

**Experiment**: A humanoid robot supports the participant of the study to solve a math puzzle. Do participants perceive the robots differently if the robots are described with anthropomorphic framing or functional framing?



---

Response variable: Perception of the robot (humanness, eeriness, acceptance)

.pull-left[

**Anthropomorphic framing**

- the robot has a name
- has a personal story
- has a favorite color and hobbies
- pronoun: him
]


.pull-right[

**Functional framing**

- height, weight
- pronoun: it
]

---

Response variable: Perception of the robot (humanness, eeriness, acceptance)

.pull-left[

**Anthropomorphic framing**

- the robot has a name
- has a personal story
- has a favorite color and hobbies
- pronoun: him




$n_1=20$  
$\bar x_1 = 3.18$  
$s_1 = 0.57$  
]


.pull-right[

**Functional framing**

- height, weight
- pronoun: it

<br>


$n_2=20$  
$\bar x_2 = 3.07$  
$s_2 = 0.29$ 
]


---
class: middle

## Conditions

- Independence within groups 

- Independence between groups

- Normality

---

class: center middle

## CI


$\bar x_1 - \bar x_2 \pm t^*_{df} \times \sqrt{\frac{s_1^2}{n_1} + \frac{s_2^2}{n_2}}$

---

## 95% CI

.pull-left[
```{r echo = TRUE}
xbar1 <- 3.18
xbar2 <- 3.07
point_estimate <- xbar1-xbar2
point_estimate

s1 <- 0.57
s2 <- 0.29
n1 <- 20
n2 <- 20


```
]

.pull-right[



```{r echo = TRUE}
critical_value <- pt(0.975, df = 19) # we use the smaller groups n-1

critical_value

se <- sqrt((s1^2)/n1 + (s2^2)/n2)
se

```

]


---

## 95% CI

```{r echo= TRUE}
point_estimate - critical_value*se
point_estimate + critical_value*se

```

95% CI = (`r point_estimate - critical_value*se` , `r point_estimate + critical_value*se`)