---
title: "Central Limit Theorem"
author: "Dr. Dogucu"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      
---

layout: true
  
<div class="my-header"></div>

<div class="my-footer"> 
 Copyright &copy; <a href="https://mdogucu.ics.uci.edu">Dr. Mine Dogucu</a>. <a href="https://creativecommons.org/licenses/by-nc-sa/4.0/">CC BY-NC-SA 4.0</a></div> 


---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = "center")
options(scipen=0)
library(tidyverse)
library(fivethirtyeight)
library(scales)

set.seed(84735)

```



class: middle

## Data

```{r message = FALSE, warning = FALSE}
lapd <- read_csv("data/lapd_2018.csv") %>% 
  rename(base_pay = `Base Pay`) %>% 
  select(base_pay)
```

We will be using payroll data from Los Angeles Police Department (LAPD) from 2018.


```{r echo = TRUE}
glimpse(lapd)
```

---

## Population Distribution

```{r}
lapd %>% 
  ggplot(aes(x = base_pay)) +
  geom_histogram(binwidth = 5000) +
  scale_x_continuous(labels = comma) +
  theme(text = element_text(size = 20)) 
  
```

---
class: middle 

## Population Mean


We have data on everyone who worked for LAPD in the year 2018. So the distribution we just looked at is a population distribution. We can go ahead and calculate the population mean ( $\mu$ ).

```{r echo = TRUE}
summarize(lapd, mean(base_pay))
```

---
class: middle 

## Population Standard Deviation

We can calculate the population standard deviation ( $\sigma$ ).

```{r cho = TRUE}
summarize(lapd, sd(base_pay))
```

---

class: middle

What if we did not have access to all this data? What would we do?

--

Rely on a sample!

---

class: middle

```{r warning = FALSE}
options(scipen = 0)
pop_mean <-  mean(lapd$`Base Pay`)
```

Let's assume we went ahead and took a (random) sample of LAPD staff and asked their salary information (and they report to us truthfully) and calculated a mean, would we find a mean of 85149.05? Why, why not?

---

class: middle

Let's pretend we have never seen the data and we do not know the population parameter $\mu$. In fact this is usually what happens in real life. We do not have the population information but we do want to know a population __parameter__ (does not necessarily have to be the mean).

--

If we took a sample and calculated the sample mean, we would name this __point estimate__ of the parameter. 


---
class: middle center



|                               | Parameter of Interest | Point Estimate / Sample Statistic |
|-------------------------------|-----------------------|-----------------------------------|
| Mean                          | $\mu$                 | $\bar x$                          |
| Difference of Two Means       | $\mu_1 - \mu_2$       | $\bar x_1 - \bar x_2$             |
| Proportion                    | $\pi$                 | $p$                               |
| Difference of Two Proportions | $\pi_1 - \pi_2$       | $p_1 - p_2$                       |


---
class: middle 

## First Sample

We would like to know about $\mu$ but we cannot access the whole population.

A researcher takes a random sample of 20 LAPD staff and ask them about their base pay.

--

```{r}
sample_1 <- lapd %>% 
  sample_n(20) 
sample_1$base_pay
```

--

**Mean of first sample**, $\bar x_1$ = 

```{r}
mean(sample_1$base_pay)
```


---

## Mean of second sample

$\bar x_2$ = 

```{r}

sample_2 <- lapd %>% 
  sample_n(20) %>% 
  select(base_pay)

mean(sample_2$base_pay)
```

--

## Mean of third sample

$\bar x_3$ = 

```{r}

sample_3 <- lapd %>% 
  sample_n(20) %>% 
  select(base_pay)

mean(sample_3$base_pay)
```

---

class: middle

We could do this over and over again. Don't you worry! I did it.

I have taken 10,000 samples of size 200 (sample size of 20 is just too small) and calculated their mean. The following slide shows the distribution of the **sample means**.

---

### Sampling Distribution of the Mean

```{r warning = FALSE, cache = TRUE}
set.seed(123)
sample_means <-vector()

for (i in 1:10000){
    sample_means <- c(sample_means, mean(sample_n(lapd, 200)$base_pay))
}

sample_means_data <- data.frame(sample_means = sample_means)

```

```{r warning = FALSE, message = FALSE}
sample_means_data %>% 
  ggplot(aes(x = sample_means)) +
  geom_histogram() +
  labs(x = "Sample Means") +
  scale_x_continuous(labels = comma) +
  theme(text = element_text(size = 20))  +
    geom_text(x = 0, y = 0, label = "bar(x)", parse=T)
```

---

### Sampling Distribution of the Mean

```{r warning = FALSE, message = FALSE}
population_mean <- mean(lapd$base_pay)

sample_means_data %>% 
  ggplot(aes(x = sample_means)) +
  geom_histogram() +
  labs(x = "Sample Means") +
  scale_x_continuous(labels = comma) +
  theme(text = element_text(size = 20))  +
    geom_text(x = 0, y = 0, label = "bar(x)", parse=T) +
  geom_vline(xintercept = population_mean)

```

---

class: middle

## Conclusion

When certain conditions are met then:


$$\bar x \sim \text{approximately }N( \mu, \frac{\sigma^2}{n})$$

--

$$(\bar x_1 - \bar x_2) \sim \text{approximately } N(\mu_1 - \mu_2, \frac{\sigma_1^2}{n_1}+ \frac{\sigma_2^2}{n_2})$$

--

$$p \sim  \text{approximately } N(\pi,  \frac{\pi(1-\pi)}{n})$$

--

$$(p_1 - p_2) \sim \text{approximately } N((\pi_1 - \pi_2),  {\frac{\pi_1(1-\pi_1)}{n_1} + \frac{\pi_2(1-\pi_2)}{n_2}})$$

---

class: middle 

## Central Limit Theorem (CLT)

If certain conditions are met, the sampling distribution will be normally distributed with a mean equal to the population parameter. The standard deviation will be inversely proportional to the square root of the sample size. 

--

We will learn the conditions in the upcoming lectures.

--

Moving forward we will use CLT to make __inference__ about population parameters using sample statistics. 

