---
title: "Visual Summaries"
author: "Dr. Dogucu"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      
---

layout: true
  
<div class="my-header"></div>

<div class="my-footer"> 
 Copyright &copy; <a href="https://mdogucu.ics.uci.edu">Dr. Mine Dogucu</a>. <a href="https://creativecommons.org/licenses/by-nc-sa/4.0/">CC BY-NC-SA 4.0</a></div> 

---


```{r message = FALSE, echo = FALSE, warning = FALSE, message=FALSE}
library(tidyverse)
library(fivethirtyeight)
library(emojifont)
library(openintro)
library(fGarch)
library(gganimate)
theme_set(theme_gray(base_size = 22))
```



## Goals

We will learn to 
- make visuals
- read and interpret visuals
- use vocabulary to describe visuals.

---

## Bar plot with counts

```{r echo = FALSE, fig.align='center'}
ggplot(candy_rankings) +
  aes(x = chocolate) +
  geom_bar()
```

---

## Bar plot with proportions

```{r echo = FALSE, fig.align='center'}
ggplot(candy_rankings) +
  aes(x = chocolate, y = ..prop.., group = 1) +
  geom_bar()
```


---

## Histogram

```{r echo = FALSE, fig.align='center'}
ggplot(candy_rankings) +
  aes(x = winpercent) +
  geom_histogram(binwidth = 3)  
```

---

## Fun fact

__histo__ comes from the Greek word _histos_ that literally means "anything set up right".

__gram__: comes from the Greek word _gramma_ which means "that which is drawn".

.footnote[Online Etymology Dictionary]


---

## Histogram

- Consider the height distribution in our class.

- How would the distribution change if Michael Jordan (198.1 cm) were to join our class?

- How would the distribution change if Tyrion Lannister (135 cm) were to join our class?

---

## Note

When data display a skewed distribution we rely on median rather than the mean to understand the center of the distribution.
---

## Histogram

---



## Box plot



---

## Histogram vs. Boxplot

.pull-left[

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.height=6, cache = TRUE}
set.seed(12345)

symmetric <- rnorm(n = 10000, mean = 0, sd = 1)

right_skewed <- rsnorm(n = 10000, mean = 0, sd = 1, xi = 20)

left_skewed <- rsnorm(n = 10000, mean = 0, sd = 1, xi = -20)

type <- c(
  rep("left-skewed", 10000),
  rep("symmetric", 10000),
  rep("right-skewed", 10000)
)

x <- c(left_skewed,
       symmetric,
       right_skewed)

data <- tibble(x = x,
                  type = type)


data %>% 
  ggplot(aes(x = x)) +
  geom_histogram() +
  transition_states(type,
                    transition_length = 1,
                    state_length = 10) +
  labs(title = 'Distribution: {closest_state}',
       x = 'observations')

```
]

.pull-right[
```{r echo = FALSE, fig.height=6, cache = TRUE, message = FALSE}
data %>% 
  ggplot(aes(x = 1, y = x)) +
  geom_boxplot() +
    theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  transition_states(type,
                    transition_length = 1,
                    state_length = 10) +
  labs(title = 'Distribution: {closest_state}',
       y = "observations")
```

]
---

## Summary



<table align = "center">
  <tr>
    <th>variable</th>
    <th>plot</th>
  </tr>
  
   <tr> 
    <td> categorical </td>
    <td> bar plot </td>
  </tr>
  
  <tr> 
  <td> numeric </td>
  <td> histogram </td>
  </tr>
  
  
  <tr> 
  <td> numeric </td>
  <td> box plot </td>
  </tr>
  
</table>

---

## Data Visualization in R

For data visualization in R we will be using the `ggplot()` function.
---

## Fun facts

- `ggplot`'s gg stands for Grammar of Graphics.
- Grammar of Graphics is a book by Leland Wilkenson.


---

If you wanted to make the following plot and R could understand English what would you tell R? 

```{r echo = FALSE, fig.height=6, fig.width=9, fig.align='center'}
ggplot(candy_rankings) +
  aes(x = chocolate) +
  geom_bar()

```

---
## Three Steps of Making a basic ggplot

- Make space for a plot using the dataset. 
- Decide on what goes on the axes.
- Decide what kind of a plot.
---

## Step 1 - Make space for the plot

.left-panel[
```{r eval = FALSE}
ggplot(candy_rankings)
```
]
.right-panel[

```{r echo = FALSE, fig.height=6, fig.width=9}
ggplot(candy_rankings)

```
]

---

## Step 2 - Decide on variables and their axes

.left-panel[
```{r eval = FALSE}
ggplot(candy_rankings) +
  aes(x = chocolate)
```
]
.right-panel[

```{r echo = FALSE, fig.height=6, fig.width=9}
ggplot(candy_rankings) +
  aes(x = chocolate)

```
]

---

## Step 3 - Decide on the type of plot

.left-panel[
```{r eval = FALSE}
ggplot(candy_rankings) +
  aes(x = chocolate) +
  geom_bar() 
```
]
.right-panel[

```{r echo = FALSE, fig.height=6, fig.width=9}
ggplot(candy_rankings) +
  aes(x = chocolate) +
  geom_bar()

```
]

---

## Practice

```{r echo = FALSE, fig.height=6, fig.width=9, fig.align='center', warning=FALSE}
ggplot(candy_rankings) +
  aes(x = winpercent) +
  geom_histogram() 

```

---

## Warning & Message

`r emoji(search_emoji('warning'))`When you see __warning__ or a __message__. The code runs and returns a ggplot but also lets us know that something is going on.
---
## What is binwidth?

Warnings and messages are mostly function specific and require some thinking (and possibly Googling) to figure things out. 

If bar plots have bars and they are called bar plots, technically histograms are bin plots because they have bins. 

The warning is basically warning us that by default, `geom_histogram()` displays 30 bins. It is asking us to set the binwidth. 

Where do you think the binwidth argument would go?

---

## Binwidth

.left-panel[
```{r eval = FALSE}
ggplot(candy_rankings) +
  aes(x = winpercent) +
  geom_histogram(binwidth = 3) 
```
]
.right-panel[

```{r echo = FALSE, fig.height=6, fig.width=9}
ggplot(candy_rankings) +
  aes(x = winpercent) +
  geom_histogram(binwidth = 3) 

```
]



---
class: center middle 

## Colors

.large[
`r emoji(search_emoji('rainbow'))[1]`
] 

Pick your favorite color(s) from the list at:

[bit.ly/colors-r](http:/bit.ly/colors-r)

---

# color = ?

.left-panel[


```{r fig.height=6, fig.width=9, eval = FALSE}
ggplot(candy_rankings) +
  aes(x = chocolate) +
  geom_bar(color = 'darkblue') 

```

]

.right-panel[
```{r echo = FALSE, fig.height=6, fig.width=9}
ggplot(candy_rankings) +
  aes(x = chocolate) +
  geom_bar(color = 'darkblue') 
```

]

---

# fill = ?

.left-panel[


```{r fig.height=6, fig.width=9, eval = FALSE}
ggplot(candy_rankings) +
  aes(x = chocolate) +
  geom_bar(color = 'darkblue',
           fill = 'coral') 

```

]

.right-panel[
```{r echo = FALSE, fig.height=6, fig.width=9}
ggplot(candy_rankings) +
  aes(x = chocolate) +
  geom_bar(color = 'darkblue',
           fill = 'coral') 
```

]



---

## Fun fact

- [BBC](https://bbc.github.io/rcookbook/)   
- [fivethirtyeight](https://blog.revolutionanalytics.com/2016/07/data-journalism-with-r-at-538.html?utm_content=bufferc03e0&utm_medium=social&utm_source=twitter.com&utm_campaign=buffer)  

---

class: center middle

[RStudio Cheat Sheets](https://rstudio.com/resources/cheatsheets/)






