---
title: "Poisson Distribution"
author: "Dr. Dogucu"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      
---

layout: true
  
<div class="my-header"></div>

<div class="my-footer"> 
 Copyright &copy; <a href="https://mdogucu.ics.uci.edu">Dr. Mine Dogucu</a>. <a href="https://creativecommons.org/licenses/by-nc-sa/4.0/">CC BY-NC-SA 4.0</a></div> 


---

## Poisson Distribution

Let $X$ represent the number of occurrences of an event with in a  __fixed__ time or space.

--

$X \sim Poisson (\lambda)$

--

$P(X = x) = f(x) =\frac{\lambda^x}{x!} e^{-\lambda}$ 

--

$S = \{0,1,2....\}$

--

$E(X) = Var(X) = \lambda$

---

## Probability Mass Function 

```{r echo = FALSE, message = FALSE, fig.align="center", fig.width=10}
library(tidyverse)
theme_set(theme_gray(base_size = 22))

x <- rep(seq(0, 20), 3)
lambda <- c(rep(1, 21), rep (5, 21), rep(10, 21))


y <- dpois(x, lambda)

data <- data.frame(x = x,
                   y = y,
                   lambda = lambda)

data$lambda <- 
  factor(data$lambda,
         levels = c("1", "5", "10"),
         labels =
           c(expression(paste(lambda, " = 1")),
             expression(paste(lambda, " = 5")),
             expression(paste(lambda, " = 10"))))

data %>% 
  ggplot() +
  aes(x = x, y = y) +
  geom_segment(data = data, 
               aes(x = x, y = 0, 
                   xend = x, yend = y)) +
  facet_wrap(data$lambda, 
             labeller = label_parsed) +
  labs(x = expression(paste("x")),
       y = expression(paste("f(X = x)")))

```

---

class:center middle

## Pangolins


<iframe width="560" height="315" src="https://www.youtube.com/embed/5BgDCDVZk8A" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>


---

class: middle

## Pangolins


It is estimated that one pangolin is snatched from the wild every five minutes ([TRAFFIC](https://www.traffic.org/what-we-do/species/pangolins/)).

---

What is the probability that 7 pangolins will be snatched from the wild in the next hour? (Note: fixed time here is an hour)

--

$P(X = 7) = f(7) = ?$

--

$E(X) = \lambda = 1 \times 12 = 12$

--

$P(X = 7) = f(7) = \frac{12^7}{7!} e^{-12} = 0.04368219$

```{r}
((12^7)/factorial(7))*exp(-12)
```

--

```{r}
dpois(x = 7, lambda = 12)
```

---

class:center middle

### pmf when $\lambda = 12$


```{r echo = FALSE, message = FALSE, fig.align="center", fig.height = 5}

x <- seq(0,30)
y <- dpois(x, 12)

data <- data.frame(x = x,
                   y = y)

data %>% 
  ggplot() +
  aes(x = x, y = y) +
  geom_segment(data = data, 
               aes(x = x, y = 0, 
                   xend = x, yend = y)) +
  labs(x = expression(paste("x")),
       y = expression(paste("f(x)")))

```

---

## cdf 

What is the probability that less than 3 pangolins will be snatched from the wild in the next hour?

--

$P(X < 3) = P(X \leq 2) = P(X = 0) + P(X = 1) + P(X = 2)$

```{r}
dpois(x = 0, lambda = 12) + 
  dpois(x = 1, lambda = 12) +
  dpois(x = 2, lambda = 12)



```

--

```{r}
ppois(q = 2, lambda = 12)
```


