---
title: "Discrete Random Variables"
author: "Dr. Dogucu"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      
---

layout: true
  
<div class="my-header"></div>

<div class="my-footer"> 
 Copyright &copy; <a href="https://mdogucu.ics.uci.edu">Dr. Mine Dogucu</a>. <a href="https://creativecommons.org/licenses/by-nc-sa/4.0/">CC BY-NC-SA 4.0</a></div> 

---

## Discrete Random Variables

A numeric outcome of a random process is called a random variable.

--

A **discrete random** variable has a countable number of possible outcomes.

--

We will use capital letters towards the end of the alphabet such as $X, Y, Z$ to represent random variables.

We will use the corresponding lower case letters $x, y, z$ to represent observed outcomes.

---

## Discrete Random Variable

Random process: 3 flips of a single coin

--

Sample Space: $\{HHH, HHT, HTH, THH, HTT, THT, TTH, TTT\}$

--

Let X be a discrete random variable that represents the total number of heads in 3 coin flips.

---

## Discrete Random Variable


<table align = "center">
<tr>
<td>x </td>

<td>0</td> <td>1</td> <td>2</td> <td>3</td>

</tr>

<tr>
<td>P(X=x) = f(x)</td>

<td>0.125</td> <td>0.375</td> <td>0.375</td> <td>0.125</td>

</tr>

</table>

--

$P(X = 2) =f(2)=\frac{3}{8} = 0.375$

---
## Probability Mass Function

For discrete random variable, $X$ the distribution of all possible values of $x$ can be shown with a probability mass function (pmf). 

```{r echo = FALSE, message = FALSE, fig.align="center", fig.height = 5}
library(tidyverse)
theme_set(theme_gray(base_size = 22))

x <- c(0, 1, 2, 3)
y <- c(0.125, 0.375, 0.375, 0.125)

data <- data.frame(x = x,
                   y = y)

data %>% 
  ggplot() +
  aes(x = x, y = y) +
  geom_segment(data = data, 
               aes(x = x, y = 0, 
                   xend = x, yend = y)) +
  labs(x = expression(paste("x")),
       y = expression(paste("f(x)")))

```

---

## Probability Mass Function (pmf)

<table align = "center">
<tr>
<td>x</td>

<td>0</td> <td>1</td> <td>2</td> <td>3</td>

</tr>

<tr>
<td>P(X=x) = f(x)</td>

<td>0.125</td> <td>0.375</td> <td>0.375</td> <td>0.125</td>

</tr>

</table>

$P(X = 2) =f(2)=\frac{3}{8} = 0.375$

--
## Cumulative Distribution Function (cdf)


<table align = "center">
<tr>
<td>x</td>

<td>0</td> <td>1</td> <td>2</td> <td>3</td>

</tr>

<tr>
<td>P(X &le;x) = F(x)</td>

<td>0.125</td> <td>0.5</td> <td>0.875</td> <td>1</td>

</tr>

</table>


$P(X \leq 2) =F(2)=\frac{3}{8} = 0.875$


---

## Cumulative Distribution Function
 

```{r echo = FALSE, message = FALSE, fig.align="center", fig.height = 5}
library(tidyverse)
theme_set(theme_gray(base_size = 22))

x <- c(0, 1, 2, 3)
y <- c(0.125, 0.5, 0.875, 1)

data <- data.frame(x = x,
                   y = y)

data %>% 
  ggplot() +
  aes(x = x, y = y) +
  geom_segment(data = data, 
               aes(x = x, y = 0, 
                   xend = x, yend = y)) +
  labs(x = expression(paste("x")),
       y = expression(paste("F(x)")))

```

---

## Expected Value

If we repeat this random process of 3 coin flips many times, how many number of heads would we expect to see on average?

<table align = "center">
<tr>
<td>x </td>

<td>0</td> <td>1</td> <td>2</td> <td>3</td>

</tr>

<tr>
<td>P(X=x) = f(x)</td>

<td>0.125</td> <td>0.375</td> <td>0.375</td> <td>0.125</td>

</tr>

</table>

$E(X) = \sum_{S} x f(x)$

--

$E(X) = (0\cdot 0.125) + (1\cdot 0.375) + (2\cdot 0.375) + (3\cdot 0.125)$

$E(X) = \mu = 1.5$

---

## Variance

Variance is the average squared deviation from the average.

---

## Variance

~~Variance is the average squared deviation from the average.~~

Variance is the expected squared deviation from the expected value.

$$Var(X)= E[(X-E(X))^2]$$
--

$$Var(X)= E[(X-\mu)^2]$$
--

$$Var(X)= E(X^2) - [E(X)]^2$$
---

## Variance


<table align = "center">
<tr>
<td>x </td>

<td>0</td> <td>1</td> <td>2</td> <td>3</td>

</tr>

<tr>
<td>P(X=x) = f(x)</td>

<td>0.125</td> <td>0.375</td> <td>0.375</td> <td>0.125</td>

</tr>

</table>

$Var(X)= E(X^2) - [E(X)]^2$

--

$E(X^2) = \sum_{S} x^2 f(x)$ 

--

$E(X^2) = (0^2\cdot 0.125) + (1^2\cdot 0.375) + (2^2\cdot 0.375) + (3^2\cdot 0.125) = 3$

--

$Var(X)= 3 - 1.5^2 = 0.75$

---

## Linear Combination of Random Variables

Let $X$ and $Y$ represent random variables and $a$ and $b$ represent constants. Then 

$E(aX +bY) = a\cdot E(X) + b \cdot E(Y)$
$Var(aX +bY) = a^2\cdot Var(X) + b^2 \cdot Var(Y)$







