---
title: "Normal Distribution"
author: "Dr. Dogucu"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      
---

layout: true
  
<div class="my-header"></div>

<div class="my-footer"> 
 Copyright &copy; <a href="https://mdogucu.ics.uci.edu">Dr. Mine Dogucu</a>. <a href="https://creativecommons.org/licenses/by-nc-sa/4.0/">CC BY-NC-SA 4.0</a></div> 


---

```{r message = FALSE, echo = FALSE}
library(tidyverse)
library(gridExtra)

theme_set(theme_gray(base_size = 22))


plot_norm_x_prob <- function(value, mean, sd, end = 3, lower = TRUE){
  
  x <- seq(mean-(end*sd),mean+(end*sd), 0.01)
  
  y <- dnorm(x, mean = mean, sd = sd)
  
  data <- data.frame(x = x, y = y)
  
  data_line <- data.frame(x = value, y = dnorm(value, mean, sd), 
                          label = paste("x = ", value))
  
  if (lower == TRUE){
    p <- qplot(x, y, data = data, 
          geom = "line") +
      geom_ribbon(data = subset(data, x > (mean-(end*sd)) & x < value),
                  aes(ymax = y), 
                  ymin = 0,
                  fill = "#e56646", 
                  colour = NA, 
                  alpha = 0.5) +
      ylab("") 

  }
  
  if (lower == FALSE){
    p <- qplot(x, y, data = data, 
          geom = "line") +
      geom_ribbon(data = subset(data, x < (mean+(end*sd)) & x > value),
                  aes(ymax = y), 
                  ymin = 0,
                  fill = "#e56646", 
                  colour = NA, 
                  alpha = 0.5) +
      ylab("")
  }
  
p
}

plot_norm_x <- function(value, mean, sd, end = 3){
  data_norm <-  data.frame(x = c(mean-(end*sd), mean+(end*sd)))
  
  data_line <- data.frame(x = value, y = dnorm(value, mean, sd), 
                          label = paste("x = ", value))
  ggplot(data_norm, 
         aes(x)) +
    stat_function(fun = dnorm, n = 1001, args = list(mean = mean, sd = sd)) + 
    ylab("") +
    geom_segment(data = data_line, aes(x = value, y = 0, xend = x, yend = y), color = "#e56646") +
    geom_text(data = data_line, aes(x = x, y = 0, label = label), color = "#e56646") +
    labs(title =  bquote("N("*mu == .(mean) ~ "," ~ sigma == .(sd)*")"))
}



plot_norm_z <- function(value, mean, sd, end = 3){
  
  data_norm <-  data.frame(x = c(mean-(end*sd), mean+(end*sd)))
  data_z <- data.frame(x = c(-end,end))

  
  data_line_z <- data.frame(z = (value-mean)/sd, y = dnorm((value-mean)/sd), 
                            label = paste("z = ", (value-mean)/sd))
  
ggplot(data_z, 
               aes(x)) +
    stat_function(fun = dnorm, n = 1001, args = list(mean = 0, sd = 1)) + 
    xlab("z")+
    ylab("") +
    geom_segment(data = data_line_z, aes(x = z, y = 0, xend = z, yend = y), color = "#e56646") +
    geom_text(data = data_line_z, aes(x = z, y = 0, label = label), color = "#e56646") +
  labs(title =  bquote("N("*mu == 0 ~ "," ~ sigma == 1*")"))

  
}
```


## pdf

$X \sim \mathcal{N}(\mu, \sigma^2)$

--

$f(x) = \frac{1}{\sigma \sqrt{2\pi}}{e^{-(x-\mu)^2/2\sigma^2}}$

--

$E(X) = \mu$

--

$Var(X) = \sigma^2$


---

class:middle 

## Characteristics

.pull-left[

```{r echo = FALSE, fig.height=4, fig.align='center'}
plot_norm_x_prob(100, 100, 15) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

```
]

.pull-right[
Mean = Median = Mode   

$P(X<\mu) = 0.5$

]
---

## Parameters


```{r echo = FALSE, fig.align='center'}
p1 <- 
  ggplot(data = data.frame(x = c(0, 200)), 
         aes(x)) +
  stat_function(fun = dnorm, n = 1001, args = list(mean = 100, sd = 15)) + 
  ylab("") +
  labs(title = "N(100, 225)")

p2 <- 
  ggplot(data = data.frame(x = c(0, 200)), 
         aes(x)) +
  stat_function(fun = dnorm, n = 1001, args = list(mean = 100, sd = 30)) + 
  ylab("") +
  labs(title = "N(100, 900)")

p3 <- 
  ggplot(data = data.frame(x = c(0, 200)), 
         aes(x)) +
  stat_function(fun = dnorm, n = 1001, args = list(mean = 90, sd = 15)) + 
  ylab("") +
  labs(title = "N(90, 225)")


p4 <- 
  ggplot(data = data.frame(x = c(0, 200)), 
         aes(x)) +
  stat_function(fun = dnorm, n = 1001, args = list(mean = 90, sd = 30)) + 
  ylab("") +
  labs(title = "N(90, 900)")

grid.arrange(p1, p2, p3, p4, ncol = 2, nrow = 2)
```

---

68.26895% of the data are within 1 standard deviation of the mean.


```{r fig.align= 'center', echo = FALSE}
mean <- 100
sd <- 15
end <- 3.5

data_line1 <- data.frame(x = mean-sd, y = dnorm(mean-sd, mean, sd), 
                          label = paste("\U03bc", "-","\U03c3" ))
data_line2 <- data.frame(x = mean+sd, y = dnorm(mean+sd, mean, sd), 
                          label = paste("\U03bc", "+","\U03c3" ))

x <- seq(mean-(end*sd),mean+(end*sd), 0.01)
  
  y <- dnorm(x, mean = mean, sd = sd)
  
  data <- data.frame(x = x, y = y)
  

  

qplot(x, y, data = data, 
          geom = "line") +
      geom_ribbon(data = subset(data, x > (mean-(1*sd)) & x < (mean+(1*sd))),
                  aes(ymax = y), 
                  ymin = 0,
                  fill = "#e56646", 
                  colour = NA, 
                  alpha = 0.5) +
      ylab("") +
    geom_text(data = data_line1, aes(x = x, y = -0.001, label = label), color = "#e56646") +
    geom_text(data = data_line2, aes(x = x, y = -0.001, label = label), color = "#e56646") +
    theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),

)




```



---


95.44997% of the data are within 2 standard deviations of the mean.


```{r fig.align= 'center', echo = FALSE}
mean <- 100
sd <- 15
end <- 3.5

data_line1 <- data.frame(x = mean-2*sd, y = dnorm(mean-(2*sd), mean, (2*sd)), 
                          label = paste("\U03bc", "-2","\U03c3" ))
data_line2 <- data.frame(x = mean+2*sd, y = dnorm(mean+(2*sd), mean, sd), 
                          label = paste("\U03bc", "+2","\U03c3" ))

x <- seq(mean-(end*sd),mean+(end*sd), 0.01)
  
y <- dnorm(x, mean = mean, sd = sd)
  
data <- data.frame(x = x, y = y)
  

  

qplot(x, y, data = data, 
          geom = "line") +
      geom_ribbon(data = subset(data, x > (mean-(2*sd)) & x < (mean+(2*sd))),
                  aes(ymax = y), 
                  ymin = 0,
                  fill = "#e56646", 
                  colour = NA, 
                  alpha = 0.5) +
      ylab("") +
    geom_text(data = data_line1, aes(x = x, y = -0.001, label = label), color = "#e56646") +
    geom_text(data = data_line2, aes(x = x, y = -0.001, label = label), color = "#e56646") +
    theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),

)




```


---

99.73002% of the data are within 3 standard deviations of the mean.

```{r fig.align= 'center', echo = FALSE}
mean <- 100
sd <- 15
end <- 3.5

data_line1 <- data.frame(x = mean-3*sd, y = dnorm(mean-(3*sd), mean, (3*sd)), 
                          label = paste("\U03bc", "-3","\U03c3" ))
data_line2 <- data.frame(x = mean+3*sd, y = dnorm(mean+(3*sd), mean, sd), 
                          label = paste("\U03bc", "+3","\U03c3" ))

x <- seq(mean-(end*sd),mean+(end*sd), 0.01)
  
y <- dnorm(x, mean = mean, sd = sd)
  
data <- data.frame(x = x, y = y)
  

  

qplot(x, y, data = data, 
          geom = "line") +
      geom_ribbon(data = subset(data, x > (mean-(3*sd)) & x < (mean+(3*sd))),
                  aes(ymax = y), 
                  ymin = 0,
                  fill = "#e56646", 
                  colour = NA, 
                  alpha = 0.5) +
      ylab("") +
    geom_text(data = data_line1, aes(x = x, y = -0.001, label = label), color = "#e56646") +
    geom_text(data = data_line2, aes(x = x, y = -0.001, label = label), color = "#e56646") +
    theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),

)




```

---

### Example 

Midterm exam scores are normally distributed with a mean of 80 and standard deviation of 5 points.

--

.pull-right[

.center[
$X \sim N(80, 5^2)$
]

```{r echo = FALSE, message = FALSE, fig.align="center", fig.height = 4, message = FALSE, warning = FALSE}

theme_set(theme_gray(base_size = 22))


ggplot(data = data.frame(x = c(65, 95)), aes(x)) +
  stat_function(fun = dnorm, 
                n = 101, 
                args = list(mean = 80, sd = 5)) +
  labs(y = "f(x)")

```


]

--

$f(78)$ =



```{r}
dnorm(x = 78, mean = 80, sd = 5)
```

--

$f(92)$ =


```{r}
dnorm(x = 92, mean = 80, sd = 5)
```

---

## Calculating Probability

What is the probability that a student scores below 73 points?


.pull-right[
```{r echo = FALSE, warning = FALSE, fig.height = 4}
mean <- 80
sd <- 5
end <- 3.5

data_line1 <- data.frame(x = mean-3*sd, y = dnorm(mean-(3*sd), mean, (3*sd)), 
                          label = paste("\U03bc", "-3","\U03c3" ))
data_line2 <- data.frame(x = mean+3*sd, y = dnorm(mean+(3*sd), mean, sd), 
                          label = paste("\U03bc", "+3","\U03c3" ))

x <- seq(mean-(end*sd),mean+(end*sd), 0.01)
  
y <- dnorm(x, mean = mean, sd = sd)
  
data <- data.frame(x = x, y = y)
  
qplot(x, y, data = data, 
          geom = "line") +
      geom_ribbon(data = subset(data, x < 73),
                  aes(ymax = y), 
                  ymin = 0,
                  fill = "#e56646", 
                  colour = NA, 
                  alpha = 0.5) +
      ylab("f(x)") 

```


]

```{r}
pnorm(q = 73, mean = 80, sd = 5)
```

--

<br>

If a student scores 73 points, the student's percentile rank is `r pnorm(73, mean = 80, sd = 5)*100`%

---

## Calculating Probability

What is the probability that a student scores above 78 points?

.pull-right[
```{r echo = FALSE, warning = FALSE, fig.height = 4}
mean <- 80
sd <- 5
end <- 3.5

data_line1 <- data.frame(x = mean-3*sd, y = dnorm(mean-(3*sd), mean, (3*sd)), 
                          label = paste("\U03bc", "-3","\U03c3" ))
data_line2 <- data.frame(x = mean+3*sd, y = dnorm(mean+(3*sd), mean, sd), 
                          label = paste("\U03bc", "+3","\U03c3" ))

x <- seq(mean-(end*sd),mean+(end*sd), 0.01)
  
y <- dnorm(x, mean = mean, sd = sd)
  
data <- data.frame(x = x, y = y)
  
qplot(x, y, data = data, 
          geom = "line") +
      geom_ribbon(data = subset(data, x >78),
                  aes(ymax = y), 
                  ymin = 0,
                  fill = "#e56646", 
                  colour = NA, 
                  alpha = 0.5) +
      ylab("f(x)") 

```
]

--
```{r}
1 - pnorm(q = 78, mean = 80, sd = 5)
```

--

```{r}
pnorm(q = 78, mean = 80, sd = 5, 
      lower.tail = FALSE)
```

---

## How many standard deviations away from the mean?

Krishna scored 85 points on the midterm. How many standard deviations is Krishna's score away from the mean?

$\frac{85-80}{5} = 1$

--

Hira scored 77.5 points on the midterm. How many standard deviations is Hira's score away from the mean?

$\frac{77.5-80}{5} = -0.5$

---

class:center middle

## Z-score

$$Z = \frac{X-\mu}{\sigma} $$

---

class: middle

## Calculating z-scores

We can calculate Z-scores regardless of the distribution of the data. In other words, data do not need to be normally distributed. Z-scores only show how many standard deviations a value is away from the mean.

---

class: middle

## Interpreting z-scores

If we know that the data follow a normal distribution then we can use z-scores for probabilistic interpretations.


---

### Normal Distribution vs. Standard Normal (z) Distribution

.pull-left[

```{r echo = FALSE, fig.height=5, fig.width=5, fig.align = 'center'}
plot_norm_x(90, mean = 80, sd = 5)
```

]

.pull-right[
```{r echo = FALSE, fig.height=5, fig.width=5, fig.align = 'center'}
plot_norm_z(2, mean = 0, sd = 1)
```
]

---
class: middle 

.pull-left[

__Normal Distribution__

```{r fig.height=5, fig.width=5, fig.align = 'center'}
pnorm(q = 90, mean = 80, sd = 5)
```

]

.pull-right[

__Standard Normal (z) Distribution__

```{r fig.height=5, fig.width=5, fig.align = 'center'}
pnorm(q = 2, mean = 0, sd = 1)
```
]

---

## SAT scores - z scores

Let X represent the SAT scores and $X \sim N(1100, 200^2)$

--

What is the Z-score of a student who scores 1500?

$z = \frac{1500-1100}{200} = 2$

---

## SAT scores - probability

What is the percentile rank of a student who scores 1500?

```{r echo = FALSE, fig.align='center', fig.height=5}

plot_norm_x_prob(1500, mean = 1100, sd = 200)

```


---


## SAT scores - probability

What is the percentile rank of a student who scores 1500?


--

.pull-left[
```{r}
pnorm(q = 1500, 
      mean = 1100, 
      sd = 200)
```


If you know the order of arguments:

```{r}
pnorm(1500, 1100, 200)
```

]

--
.pull-right[

```{r}
pnorm(2, 
      mean = 0, 
      sd = 1)
```

The default `mean` is 0 and `sd` is 1.

```{r}
pnorm(2)
```
]

---

What percentage of test takers score between 800 and 1400?

--

$P(800<X<1400) =P(X<1400) - P(X<800)$

--

```{r echo = FALSE, warning = FALSE, fig.height = 3.75, fig.align='center'}
mean <- 1100
sd <- 200
end <- 3.5

data_line1 <- data.frame(x = mean-3*sd, y = dnorm(mean-(3*sd), mean, (3*sd)), 
                          label = paste("\U03bc", "-3","\U03c3" ))
data_line2 <- data.frame(x = mean+3*sd, y = dnorm(mean+(3*sd), mean, sd), 
                          label = paste("\U03bc", "+3","\U03c3" ))

x <- seq(mean-(end*sd),mean+(end*sd), 0.01)
  
y <- dnorm(x, mean = mean, sd = sd)
  
data <- data.frame(x = x, y = y)
  
qplot(x, y, data = data, 
          geom = "line") +
      geom_ribbon(data = subset(data, x < 1400 & x > 800),
                  aes(ymax = y), 
                  ymin = 0,
                  fill = "#e56646", 
                  colour = NA, 
                  alpha = 0.5) +
      ylab("f(x)") 

```

--

```{r}
pnorm(q = 1400, mean = 1100, sd = 200) -
  pnorm(q = 800, mean = 1100, sd = 200) 
```


---

## Known probability unknown x

If College Board wanted to send a congratulatory email to the top 10% in the SATs, above what SAT score should they consider sending the email to?

--
.pull-left[
```{r echo = FALSE, fig.height=4, fig.align='center'}

data_text <- data.frame(x = 1356,
                        y = -0.0001,
                        label = "x = ?")

data_text2 <- data.frame(x = 1420,
                        y = 0.00025,
                        label = "10%")

plot_norm_x_prob(1356.31, 1100, 200)  +
   geom_text(data = data_text, 
             aes(x = x, 
                 y = y, 
                 label = label), 
             color = "#e56646") +
   geom_text(data = data_text2, 
             aes(x = x, 
                 y = y, 
                 label = label), 
             color = "#e56646") 
  
```
]

--

.pull-right[
```{r}
qnorm(p = 0.90, 
      mean = 1100, 
      sd = 200) 
```
]
