---
title: "Hypothesis Testing for Difference of Two Proportions"
author: "Dr. Dogucu"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      
---

layout: true
  
<div class="my-header"></div>

<div class="my-footer"> 
 Copyright &copy; <a href="https://mdogucu.ics.uci.edu">Dr. Mine Dogucu</a>. <a href="https://creativecommons.org/licenses/by-nc-sa/4.0/">CC BY-NC-SA 4.0</a></div> 


---


class: center middle

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = "center")
library(tidyverse)
options(scipen=0)
theme_set(theme_gray(base_size = 20))

set.seed(84735)


```



class: middle 

## Hypothesis Testing

Steps

1. Set hypotheses
2. Identify Sampling Distribution of $H_0$  
3. Calculate p-value  
4. Make a Decision and a Conclusion.

---

## Example

Is there a relationship between taking a college level science class and belief in life after death? Below are the responses from General Social Survey in 2018.



<table align = "center">
<thead>
  <tr>
    <th colspan="2" rowspan="2"></th>
    <th colspan="2">Belief in Life After Death</th>
  </tr>
  <tr>
    <td>Yes</td>
    <td>No<br></td>
  </tr>
</thead>
<tbody>
  <tr>
    <td rowspan="2">College Science Course<br></td>
    <td>Yes</td>
    <td>375</td>
    <td>75</td>
  </tr>
  <tr>
    <td>No</td>
    <td>485</td>
    <td>115</td>
  </tr>
</tbody>
</table>

---
class: middle center
## Step 1. Set hypotheses

$$H_0: \pi_1 = \pi_2$$

$$H_A: \pi_1 \neq \pi_2$$


---

class: middle

## Step 2. Identify Sampling Distribution of $H_0$ 

If conditions are met then according to CLT $(p_1 - p_2) \sim \text{approximately } N(\pi_1 - \pi_2,  {\frac{\pi_1(1-\pi_1)}{n_1} + \frac{\pi_2(1-\pi_2)}{n_2}})$

--

Assuming that the null is true then $$\pi_1 = \pi_2$$ so we cannot use different $p_1$ and $p_2$ in place of $\pi_1$ and $\pi_2$.


---

class: middle

## Pooled estimate of a proportion when $H_0: \pi_1 = \pi_2$

$p_{pooled} = \frac{\text{number of total successes}}{\text{number of total cases}} = \frac{p_1n_1+p_2n_2}{n_1+n_2}$

$SE = \sqrt{\frac{p_{pooled}(1-p_{pooled})}{n_1}+\frac{p_{pooled}(1-p_{pooled})}{n_2}}$

We also use the pooled proportion when checking conditions for success-failure counts.

---

class: middle

## Calculating pooled proportion

$p_{pooled} = \frac{\text{number of total successes}}{\text{number of total cases}} = \frac{p_1n_1+p_2n_2}{n_1+n_2}$

```{r echo = TRUE}
p1 <- 0.83
p2 <- 0.8083
n1 <- 450
n2 <- 600

p_pooled <- (p1*n1+p2*n2)/(n1+n2)

p_pooled
```

---

## Calculating standard error


$SE = \sqrt{\frac{p_{pooled}(1-p_{pooled})}{n_1}+\frac{p_{pooled}(1-p_{pooled})}{n_2}}$

```{r echo = TRUE}
sqrt((p_pooled*(1-p_pooled)/n1) + (p_pooled*(1-p_pooled)/n2))
```

---

class: middle

## Step 3. Calculate p-value  

.pull-left[

```{r echo = TRUE}
p1 - p2
```
How likely are we to observe a difference of proportions in samples that is at least as extreme as (0.0217)?


]

.pull-right[

```{r fig.height = 4}
se <- sqrt((p_pooled*(1-p_pooled)/n1) + (p_pooled*(1-p_pooled)/n2))

diff <- seq(-0.1, 0.1, by = 0.0001)

y <- dnorm(diff, mean = 0, sd = se)

data <- tibble(diff = diff, y = y)

qplot(diff, y, data = data, 
            geom = "line") +
        ylab("") +
        geom_ribbon(data = subset(data,  diff < -0.0217),
                    aes(ymax = y), 
                    ymin = 0,
                    fill = "#e56646", 
                    colour = NA, 
                    alpha = 0.5) +
      
        geom_ribbon(data = subset(data,  diff > 0.0217),
                    aes(ymax = y), 
                    ymin = 0,
                    fill = "#e56646", 
                    colour = NA, 
                    alpha = 0.5) +
  geom_segment(x = -0.0217, 
                 xend = -0.0217, 
                 y = 0, 
                 yend = dnorm(-0.0217, mean = 0, sd = se)) + 
    geom_segment(x = 0.0217, 
                 xend = 0.0217, 
                 y = 0, 
                 yend = dnorm(0.0217, mean = 0, sd = se)) +
  scale_x_continuous(breaks = c(-0.05, -0.022, 0, 0.022, 0.05)) +
  labs(x = expression(p[1]-p[2]))

```

]

---

If the null hypothesis is true then 

```{r echo = TRUE}

se <- sqrt((p_pooled*(1-p_pooled)/n1) + (p_pooled*(1-p_pooled)/n2))

pnorm(0.0217, mean = 0, sd = se, lower.tail = FALSE) #P(p1-p2>0.0217)
```


```{r echo = TRUE}
pnorm(-0.0217, mean = 0, sd = se) #P(p1-p2<-0.0217)
```


```{r echo = TRUE}
pnorm(-0.0217, mean = 0, sd = se)*2 #p-value

```

---

class: middle

## Step 4. Make a Decision and a Conclusion.

If the null hypothesis was true ( $\pi_1 - \pi_2 = 0$  ) then the probability of observing a difference of proportions in the sample that is at least extreme as 0.0217 would be 0.37. In other words, p-value = 0.37 which is not less than 0.05. This implies that the observed value ( $p_1 - p_2 = 0.0217$ ) is not significantly different than 0. We fail to reject the null and conclude that we have not found any evidence against the null. 

