---
title: "Discrete Uniform Distribution"
author: "Dr. Dogucu"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      
---

layout: true
  
<div class="my-header"></div>

<div class="my-footer"> 
 Copyright &copy; <a href="https://mdogucu.ics.uci.edu">Dr. Mine Dogucu</a>. <a href="https://creativecommons.org/licenses/by-nc-sa/4.0/">CC BY-NC-SA 4.0</a></div> 

---

```{r message = FALSE, echo = FALSE}
library(openintro)
library(tidyverse)
library(titanic)
```










## Review

It is estimated that 45.3% of emails we receive are spam [(Statista)](https://www.statista.com/statistics/420400/spam-email-traffic-share-annual/). Mohsen has received 30 emails today. 

1) How many of these emails would you expect to be spam? Call this value $\mu$.

2) If Mohsen was to receive 30 emails every day would you expect $\mu$ many spam email every time? Is there variance? Calculate. 

3) Pause and think. Of the 30 emails he received, what is the probability that 29 are spam? 

4) What is the probability that the number of spam emails he received is less than 9?

---


---
## Discrete Uniform Distribution

In discrete uniform distribution, all values are equally likely to be observed. A fair roll die follows a uniform distribution. 

$X\sim unif(a,b)$

pmf: $P(X=x) = \frac{1}{n}$

$\Omega_X = \{a, a+1, ..., b-1, b\}$

---
## Discrete Uniform Distribution

A broken phone alarm starts ringing once every day randomly. Let X be the random variable representing the hour on the 24 hour system when the alarm goes on. 

Then $\Omega_X = \{0, 1, 2, ...,22,23\}$

$X\sim unif(0,23)$

The probability that the alarm goes sometime between 15:00 - 16:00.

$P(X = 15) = \frac{1}{24} = 0.04167$

---

## Discrete Uniform Distribution


Note that the R function `dunif()` is used to calculate probability density function of continuous uniform distributions and __not__ for probability mass function of discrete uniform distributions. 


---

## Discrete Uniform Distribution

R is an open-source language. That means any body can contribute to R. 

Luckily someone wanted to contribute with a discrete uniform probability mass function.

However, this function is not installed when R is installed.


---

## R packages

When we download R or open RStdudio cloud, we have R functions that we can use such as `dbinom()`. These already installed set of functions are part of __base R__.

The rest comes from the open source community who develop functions, datasets and provide it for the public use in R packages. 

For instance the `ddunif()` function comes from the `fitur` package. 

---
## Using R packages

In order to use a package, you first have to make sure it is downloaded. 

```{r eval = FALSE}
install.packages("fitur")
```

You only do installation process once. Once it is downloaded it will always be downloaded to your computer.

---
## Using R packages

Once a package is downloaded you can use any function or object provided within the package. In order to use the `ddunif()` function from `fitur` package we have two options. 

Option 1:
```{r message = FALSE}
library(fitur)
ddunif(x = 15, min = 0, max = 23)
```

Use this option if you are going to keep using functions from the same package over and over again. `library(fitur)` and all packages you load should be at the top of your code.

---

## Using R packages

Option 2:
```{r}
fitur::ddunif(x = 15, min = 0, max = 23)
```

Use this option if you are going to use functions from this package only few times. 

Use this option if there is a function name conflict (i.e. another function with the same name exists in another package that you are using)

Also, this is a good option if you are unfamiliar with a package and you want to remind yourself which package the function belongs to. 

---



