---
title: "Simpson's Paradox"
author: "Dr. Dogucu"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      
---

layout: true
  
<div class="my-header"></div>

<div class="my-footer"> 
 Copyright &copy; <a href="https://mdogucu.ics.uci.edu">Dr. Mine Dogucu</a>. <a href="https://creativecommons.org/licenses/by-nc-sa/4.0/">CC BY-NC-SA 4.0</a></div> 

---

```{r message = FALSE, echo = FALSE}
library(tidyverse)
library(datasets)
```

## Simpson's Paradox - UC Berkeley Admissions, 1973

```{r echo = FALSE, fig.height=5, fig.align='center', message = FALSE}
ucb <- data.frame(UCBAdmissions)

gender <-
  group_by(ucb, Gender) %>%
  summarize(Total = sum(Freq))


gender_admit <-
  group_by(ucb, Admit, Gender) %>%
  summarize(Freq = sum(Freq)) %>%
  inner_join(gender) %>%
  mutate(Prop = Freq/Total)



gender_admit %>% 
  ggplot(aes(x = Gender, y = Prop, fill = Admit)) +
  geom_bar(stat = "identity") +
  theme(text = element_text(size=30)) +
  labs(title = "Admit Rate by Gender")


```

---

## Simpson's Paradox - UC Berkeley Admissions, 1973

```{r echo = FALSE, fig.height=5, fig.align='center', message = FALSE}

dept <-
  group_by(ucb, Dept) %>%
  summarize(Total = sum(Freq))


dept_admit <-
  group_by(ucb, Admit, Dept) %>%
  summarize(Freq = sum(Freq)) %>%
  inner_join(dept) %>%
  mutate(Prop = Freq/Total)



ggplot(dept_admit, aes(x = Dept, 
                         y = Prop, fill = Admit)) +
  geom_bar(stat = "identity")   +
  theme(text = element_text(size=30)) +
  labs(title = "Admit Rate by Dept")


```


---

## Simpson's Paradox - UC Berkeley Admissions, 1973

```{r echo = FALSE, fig.height=5, fig.align='center', message = FALSE}

gender_dept <-
  group_by(ucb, Gender, Dept) %>%
  summarize(Freq = sum(Freq)) %>%
  inner_join(dept) %>%
  mutate(Prop=Freq/Total)

ggplot(gender_dept, aes(x = Dept, y = Prop, 
                        fill = Gender)) +
  geom_bar(stat = "identity") +
  theme(text = element_text(size=15)) +
  labs(title = "App Rate based on Gender for each Dept")
  
```


---

## Simpson's Paradox 

If we observe a certain trend between two variables and this effect disappears or reverses when a third variable is introduced then this phenomenon Simpson's paradox. 

More examples on [Wikipedia](https://en.wikipedia.org/wiki/Simpson%27s_paradox)

---

## Moral of the Story

We need to move beyond thinking about the relationship between just two variables. We need to keep asking if there are/could be any **confounding** variables.

